### 2009-12-08 docs/marcello-dissertation/thesis/chapter10-appendix.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

This section includes the source-code for the DSP Data Persistence. Each
section details one single main artifact used on the development of the
component. However, only the main ones will be added into this document, while
the remainder can be downloaded directly from the Subversion repository.

@@

This section focus on the development of a Software Platform for the NetBEAMS
Gateway Embedded System, or on the system developed in the Gumstix device. As
shown in figure \ref{fig:netbeams-software-stack}, the NetBEAMS Gateway Node is
an embedded system that contains a set of layer found in any computer.
However, each of the resources described from the system were customized for
the NetBEAMS system. The main components of such system can be summarized as
follows:

@@

As depicted in figure \ref{fig:netbeams-software-stack}, the DSP Sonde Data
Producer component opens a serial connection with the device through the Java
Native Interface (JNI) driver and produces the data flow 1. The data stream
format at this stage is as described on table \ref{tab:ysi-data-stream}.
Later, by parsing and transforming the data to DSP internal representation, it
sends the data to the DSP Platform as depicted on flow 2. Finally, the DSP
Platform decides to send the data to the DSP Wire Transport Client, as shown
on flow 3, by following rules defined by the system administrator. At this
point, the data is ready to be transmitted to the server, which is covered on
the following sections.

\subsection{The DSP Platform Architecture}
\label{netbeams-architecture}

@@

In order to show an example, consider the DSP Platform on the embedded system
Gumstix in figure \ref{fig:sf-netbeams-node}. It is started by providing a
deployment configuration artifact config.xml included in the appendix. It lists
the components that the DSP Platform needs to install and load for its
operation. As it is shown, the DSPWireTransportClient and the DSPSondeProducer
are the identification of the components, followed by the name of the physical
file and an assigned loading priority. While the DSPWireTransportClient is
responsible for the serialization and deserialization of the DSP Messages
exchanged by the DSPSondeProducer and any remote DSP component. In this case,
the DSPSondeProducer is the DSP component responsible to collect measurement
data from the YSI Sonde and temporarily store them before being transmitted.

@@

These DSP Classes and other can be seen in figure the  UML Class
Diagram \cite{uml} of figure \ref{fig:DSP-Messages-Classes}. Each DSP Message
contains an instance of a Header and Body. Whenever a DSP component is ready
to transmit messages, it wraps up the set of DSP Messages into the instance of
the Messages Container, which contains information about the collection of
messages being transmitted with its own identification. In this fashion, the
Messages Container is the main communication unit between 2 different DSP
Components.

@@

All in all, the DSP Platform requires the knowledge of the aforementioned DSP
classes in order to initialize. The UML Sequence Diagram \cite{uml} in figure
\ref{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
depicts the exact moments when the additional the plug-and-play DSP components
are installed and bootstrapped by the DSP Platform, as the UML State Diagram
\cite{uml} in figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}
described.

@@

In general, when a DSP Component finishes preparing the DSP Messages Container,
it contacts an entity responsible for sending and proxying message to other
components. The DSP Broker acts like a mailman, receiving messages and
forwarding copies of it to other entities in the system, abstraction in which
is shown in figure \ref{fig:DSP-Message-Broker-Abstraction}. In this way, the
DSP Broker receives a DSP Message and makes necessary deliveries to DSP
components that it has knowledge of.

@@

As described by \cite{netbeams2009}, NetBEAMS is a single-hop sensor network
whose sensors transmits data from the sensors to the server. Upon receiving the
data stream, the DSP Wire Transport Server component desirializes the DSP
MessagesContainer and its enclosing DSP Messages back to a Java class instance
and delivers each of the DSP Messages to the DSP Broker, which delivers a
copy of the message to each DSP Component listed by the maching rule. 
%This
%process is depicted by the UML Sequence Diagram \cite{uml} in figure \ref{fig:}

@@

\subsection{Execution Logs}

@@

- starting the db
- Stopping the Database

@@

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/view-collections-instance-browser-futondb}
  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
  \label{fig:view-collections-instance-browser-futondb}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/view-collected-data-list-browser-futondb}
  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
  \label{fig:view-collected-data-list-browser-futondb}
\end{figure}

By clicking in one of the items, one can view the instance of the document, as
it is shown in figure \ref{fig:view-collections-instance-browser-futondb}.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/view-collected-data-instance-browser-futondb}
  \caption{Viewing an instance of collected data using the Futon for
  CouchDB/MongoDB}
  \label{fig:view-collected-data-instance-browser-futondb}
\end{figure}

@@

@@

>>>>

This section includes the source-code for the DSP Data Persistence, as well as
the experiments artifacts used. Each section details one single main artifact
used on the development of the component. However, only the main ones will be
added into this document, while the remainder can be downloaded directly from
the Subversion repository.

@@

This section focus on the development of a Software Platform for the NetBEAMS
Gateway Embedded System, or on the system developed in the Gumstix device. As
shown in Figure \ref{fig:netbeams-software-stack}, the NetBEAMS Gateway Node is
an embedded system that contains a set of layer found in any computer.
However, each of the resources described from the system were customized for
the NetBEAMS system. The main components of such system can be summarized as
follows:

@@

As depicted in Figure \ref{fig:netbeams-software-stack}, the DSP Sonde Data
Producer component opens a serial connection with the device through the Java
Native Interface (JNI) driver and produces the data flow 1. The data stream
format at this stage is as described on table \ref{tab:ysi-data-stream}.
Later, by parsing and transforming the data to DSP internal representation, it
sends the data to the DSP Platform as depicted on flow 2. Finally, the DSP
Platform decides to send the data to the DSP Wire Transport Client, as shown
on flow 3, by following rules defined by the system administrator. At this
point, the data is ready to be transmitted to the server, which is covered on
the following sections.

\subsection{The DSP Platform Architecture}
\label{sec:netbeams-architecture}

@@

In order to show an example, consider the DSP Platform on the embedded system
Gumstix in Figure \ref{fig:sf-netbeams-node}. It is started by providing a
deployment configuration artifact config.xml included in the appendix. It lists
the components that the DSP Platform needs to install and load for its
operation. As it is shown, the DSPWireTransportClient and the DSPSondeProducer
are the identification of the components, followed by the name of the physical
file and an assigned loading priority. While the DSPWireTransportClient is
responsible for the serialization and deserialization of the DSP Messages
exchanged by the DSPSondeProducer and any remote DSP component. In this case,
the DSPSondeProducer is the DSP component responsible to collect measurement
data from the YSI Sonde and temporarily store them before being transmitted.

@@

These DSP Classes and other can be seen in Figure the  UML Class
Diagram \cite{uml} of Figure \ref{fig:DSP-Messages-Classes}. Each DSP Message
contains an instance of a Header and Body. Whenever a DSP component is ready
to transmit messages, it wraps up the set of DSP Messages into the instance of
the Messages Container, which contains information about the collection of
messages being transmitted with its own identification. In this fashion, the
Messages Container is the main communication unit between 2 different DSP
Components.

@@

All in all, the DSP Platform requires the knowledge of the aforementioned DSP
classes in order to initialize. The UML Sequence Diagram \cite{uml} in figure
\ref{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
depicts the exact moments when the additional the plug-and-play DSP components
are installed and bootstrapped by the DSP Platform, as the UML State Diagram
\cite{uml} in Figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}
described.

@@

In general, when a DSP Component finishes preparing the DSP Messages Container,
it contacts an entity responsible for sending and proxying message to other
components. The DSP Broker acts like a mailman, receiving messages and
forwarding copies of it to other entities in the system, abstraction in which
is shown in Figure \ref{fig:DSP-Message-Broker-Abstraction}. In this way, the
DSP Broker receives a DSP Message and makes necessary deliveries to DSP
components that it has knowledge of.

@@

As described by \cite{netbeams2009}, NetBEAMS is a single-hop sensor network
whose sensors transmits data from the sensors to the server. Upon receiving the
data stream, the DSP Wire Transport Server component desirializes the DSP
MessagesContainer and its enclosing DSP Messages back to a Java class instance
and delivers each of the DSP Messages to the DSP Broker, which delivers a
copy of the message to each DSP Component listed by the maching rule. 
%This
%process is depicted by the UML Sequence Diagram \cite{uml} in Figure \ref{fig:}

@@

@@

In order to start the database, just type ``mongod --dbpath DIRECTORY/

@@

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{../diagrams/view-collections-instance-browser-futondb}
  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
  \label{fig:view-collections-instance-browser-futondb}
\end{figure}

@@

\begin{figure}[h]
  \centering

  \includegraphics[scale=0.45]{../diagrams/view-collected-data-list-browser-futondb}
  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
  \label{fig:view-collected-data-list-browser-futondb}
\end{figure}

By clicking in one of the items, one can view the instance of the document, as
it is shown in Figure \ref{fig:view-collections-instance-browser-futondb}.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{../diagrams/view-collected-data-instance-browser-futondb} \caption{Viewing an instance of collected data using the Futon for CouchDB/MongoDB}
  \label{fig:view-collected-data-instance-browser-futondb}
\end{figure}

@@

\lstinputlisting[language=Bash,label=file:setup-os-shards-step1,caption=Experiment
Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step1}

\lstinputlisting[language=Bash,label=file:setup-os-shards-step2,caption=Experiment
Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step2}

@@

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.4]{../diagrams/experiment-mouseactions-dsp}
  \caption{Experiment showing the mongoDB persisting messages from the DSP
  Mouse Actions Demo}
  \label{fig:experiment-mouseactions-dsp}
\end{figure}

### 2009-12-08 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

Data management for large scales of data require different techniques to store
and retrieve data, since it requires more computational power and data
retrieval efficient algorithms. \cite{sn-data-center} describes the creation of
a Sensor Data Center designed to store and analyze amounts of data on the range of
Terabytes to a future Petabyes of data. Its sensor network deals with not only
numerical sensor readings from transportation managemenet, but also radar
images, used for historic purposes and data mining that performs traffic
pattern analysis. In this way, the authors report the use of parallel data
processing techniques in order to provide processing for the collected
data, and load-balancing when storing data to minimize the sensor data
processing delays. For this reason, they suggest the use of concurrent control
such as snapshot isolation of the current data into separated replicated data.
In this way, data access from users are not interrupted by the processing of
the data that is processed.

>>>>

Data management for large scales of data require different techniques to store
and retrieve data, since it requires more computational power and data
retrieval efficient algorithms. \cite{sn-data-center} describes the creation of
a Sensor Data Center designed to store and analyze amounts of data on the range of
Terabytes to a future Petabyes of data. Its sensor network deals with not only
numerical sensor readings from transportation managemenet, but also radar
images, used for historic purposes and data mining that performs traffic
pattern analysis. In this way, the authors report the use of parallel data
processing techniques in order to provide processing for the collected
data, and load-balancing when storing data to minimize the sensor data
processing delays. For this reason, they suggest the use of concurrent control
such as snapshot isolation of the current data into separated replicated data.
Therefore, data access from users are not interrupted by the processing of
the data that is processed.

All in all, this chapter described a subset of problems and solutions used by
the sensor networks community. First, the survey in sensor networks was
described to give a foundation of specific properties of sensor networks. Then,
different surveys in data persistence in sensor networks were reviewed,
followed by the data models and query engines. Last, but not least, different
project described what is important to describe the collected to finally verify
different properties of data load and management. The following chapter will
clssify the properties found in this chapter in terms of taxonomies.

### 2009-12-08 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

\section{The Location of the Sensor Data Taxonomy}

@@

\begin{table}[!b]
    \label{tab:taxonomies-list-summary}
    \begin{center}
        \begin{tabular}{|p{170pt}|p{250pt}|}\hline 
        \textbf{Taxonomy Name} & \textbf{Associated Taxas}\\\hline 
        Purpose of the Sensor Data & Real-time Data Stream, Data Archival \\\hline 
        The Location of the Sensor Data & Local Storage, External Storage\\\hline
        Data Model & Schema-Dependent, Schema-less Models\\\hline 
        Data Description & Provenance-Aware, Annotations\\\hline
        Query Processing & In-Network Query, Centralized Query\\\hline 
        Data Volume & Small, Medium, Large\\\hline
        System Organization & Centralized System, Distributed System\\\hline
        \end{tabular}
    \end{center}
    \caption{Summary of the Data Persistence Taxonomies}
\end{table}

@@

In the light of these taxonomies, the classification of the data persistence
properties of a sensor networks can be studied and reviewed. As described in
the introduction of this dissertation, one the main motivations for this work
was to provide means of data persistence of NetBEAMS, a case study which does
not contain a data layer. Therefore, the following chapter seeks to describe
that case study in order to make a section of a given technology for evaluation.

>>>>

\section{Location of the Sensor Data Taxonomy}

@@

\begin{table}[!b]
    \label{tab:taxonomies-list-summary}
    \begin{center}
        \begin{tabular}{|p{170pt}|p{250pt}|}\hline 
        \textbf{Taxonomy Name} & \textbf{Associated Taxas}\\\hline 
        Purpose of the Sensor Data & Real-time Data Stream, Data Archival \\\hline 
        Location of the Sensor Data & Local Storage, External Storage\\\hline
        Data Model & Schema-Dependent, Schema-less Models\\\hline 
        Data Description & Provenance-Aware, Annotations\\\hline
        Query Processing & In-Network Query, Centralized Query\\\hline 
        Data Volume & Small, Medium, Large\\\hline
        System Organization & Centralized System, Distributed System\\\hline
        \end{tabular}
    \end{center}
    \caption{Summary of the Data Persistence Taxonomies}
\end{table}

@@

In the light of these taxonomies, the classification of the data persistence
properties of a sensor networks can be studied and reviewed. As described in
the introduction of this dissertation, one the main motivations for this work
was to provide means of data persistence for NetBEAMS, a case study which does
not contain a data layer. Therefore, the following chapter seeks to describe
NetBEAMS and its supporting infrastructure.

### 2009-12-08 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

After presenting NetBEAMS and its infrastructure, the the next step is the
actual analysis of existing database systems that fits to the requirements
and characteristics of NetBEAMS. Meanwhile, more details regarding the
architecture of NetBEAMS, as well as development guidelines, can be reached in
the appendix of chapter 10.

>>>>

After presenting the SF-BEAMS sensor network, NetBEAMS was presented as the
motivation and major study for this work. In this way, the infrastructure of
NetBEAMS were described its infrastructure, the next chapter is the actual
analysis of existing database systems mentioned on the review of the
literature, that fits to the requirements and characteristics of NetBEAMS.
Meanwhile, more details regarding the architecture of NetBEAMS, as well as
development guidelines, can be found in the appendix of chapter 10.
